---
title: "Assignment 09: Data Scraping"
author: "Yunting Wang"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

# Total points:

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on data scraping. 

## Directions
1. Change "Student Name" on line 3 (above) with your name.
2. Work through the steps, **creating code and output** that fulfill each instruction.
3. Be sure to **answer the questions** in this assignment document.
4. When you have completed the assignment, **Knit** the text and code into a single PDF file.
5. After Knitting, submit the completed exercise (PDF file) to the dropbox in Sakai. Add your last name into the file name (e.g., "Fay_09_Data_Scraping.Rmd") prior to submission.


## Set up 
1. Set up your session:

* Check your working directory
* Load the packages `tidyverse`, `rvest`, and any others you end up using.
* Set your ggplot theme

```{r, message = FALSE}
#1
getwd()
library(tidyverse)
library(rvest)
library(lubridate)
mytheme <- theme_classic() +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
theme_set(mytheme)
```

2. We will be scraping data from the NC DEQs Local Water Supply Planning website, specifically the Durham's 2019 Municipal Local Water Supply Plan (LWSP): 
 * Navigate to https://www.ncwater.org/WUDC/app/LWSP/search.php
 * Change the date from 2020 to 2019 in the upper right corner.
 * Scroll down and select the LWSP link next to Durham Municipality. 
 * Note the web address: <https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2020>

Indicate this website as the as the URL to be scraped. (In other words, read the contents into an
`rvest` webpage object.)

```{r set.the.scraping.website}
#2
the.url=read_html('https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2020')

```

3. The data we want to collect are listed below:

* From the "1. System Information" section:
 * Water system name
 * PSWID
 * Ownership
 
* From the "3. Water Supply Sources" section:
 * Average Daily Use (MGD) - for each month

In the code chunk below scrape these values, assigning them to three separate variables.

>HINT: The first value should be "Durham", the second "03-32-010", the third "Municipality", and the last should be a vector of 12 numeric values, with the first value being 36.0100.

```{r scrape.the.data}
#3
water.system.name <- the.url %>% 
  html_nodes('div+ table tr:nth-child(1) td:nth-child(2)') %>%
  html_text()
water.system.name
pswid <- the.url %>% 
  html_nodes('td tr:nth-child(1) td:nth-child(5)') %>%
  html_text()
pswid
ownership <- the.url %>% 
  html_nodes('div+ table tr:nth-child(2) td:nth-child(4)') %>%
  html_text()
ownership
max.withdrawals.mgd <- the.url %>% 
  html_nodes('th~ td+ td') %>%
  html_text()
max.withdrawals.mgd
```


4. Convert your scraped data into a dataframe. This dataframe should have a column for each of the 4 variables scraped and a row for the month corresponding to the withdrawal data. Also add a Date column that includes your month and year in data format. (Feel free to add a Year column too, if you wish.)

>TIP: Use `rep()` to repeat a value when creating a dataframe.

>NOTE: It's likely you won't be able to scrape the monthly widthrawal data in order. You can overcome this by creating a month column in the same order the data are scraped: Jan, May, Sept, Feb, etc...

5. Plot the max daily withdrawals across the months for 2020

```{r create.a.dataframe.from.scraped.data}
#4
the.month=c(1,5,9,2,6,10,03,7,11,4,8,12)
the.year=rep(2020,12)
Date=paste0(the.year,"-",the.month)
the.date=my(paste(the.month,"-",the.year))

the.ownership=rep(ownership,12)
pwsid=rep(pswid,12)
ex4=data.frame(Water.System.Name=rep(water.system.name,12),
               Ownership=rep(ownership,12),
               PWSID=rep(pswid,12),
               Max.Day.Use=as.numeric(max.withdrawals.mgd),
               Date=the.date)
ex4

#5
ex5=ggplot(ex4,aes(x=Date,y=Max.Day.Use))+
  geom_line()+
  ylab("Max Daily Withdrawals")

print(ex5)

```

6. Note that the PWSID and the year appear in the web address for the page we scraped. Construct a function using your code above that can scrape data for any PWSID and year for which the NC DEQ has data. **Be sure to modify the code to reflect the year and site scraped**.

```{r construct.a.scraping.function}
#6.
scrap.it=function(the_pwsid,the_year){
  the_website= read_html(paste0('https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=',
                                the_pwsid,'&year=',the_year))

water.system.name <- the_website %>% 
  html_nodes('div+ table tr:nth-child(1) td:nth-child(2)') %>%
  html_text()


ownership <- the_website %>% 
  html_nodes('div+ table tr:nth-child(2) td:nth-child(4)') %>%
  html_text()

max.withdrawals.mgd <- the_website %>% 
  html_nodes('th~ td+ td') %>%
  html_text()

the.month=c(1,5,9,2,6,10,03,7,11,4,8,12)
the.year=rep(the_year,12)
the.date=my(paste(the.month,"-",the_year))

the_df=data.frame(Water.System.Name=rep(water.system.name,12),
               Ownership=rep(ownership,12),
               PWSID=rep(the_pwsid,12),
               Max.Day.Use=as.numeric(max.withdrawals.mgd),
               Date=the.date)
return(the_df)
  }


```

7. Use the function above to extract and plot max daily withdrawals for Durham (PWSID='03-32-010') for each month in 2015

```{r fetch.and.plot.Durham.2015.data}
#7
ex7=scrap.it('03-32-010',2015)
ex7.plot=ggplot(ex7,aes(x=Date,y=Max.Day.Use))+
  geom_line()+
  ylab("Max Daily Withdrawals")
print(ex7.plot)

```

8. Use the function above to extract data for Asheville (PWSID = 01-11-010) in 2015. Combine this data with the Durham data collected above and create a plot that compares the Asheville to Durham's water withdrawals.

```{r fetch.and.plot.Durham.2015.data.2}
#8
ex8=scrap.it('01-11-010',2015)
ex8.combined=rbind(ex7,ex8)
ex8.plot=ggplot(ex8.combined,
                aes(x=Date,
                    y=Max.Day.Use,
                    color=Water.System.Name))+
  geom_line()+
  ylab("Max Daily Withdrawals")
print(ex8.plot)
```


9. Use the code & function you created above to plot Asheville's max daily withdrawal by months for the years 2010 thru 2019.Add a smoothed line to the plot.

```{r}
#9
the_years=seq(2010,2019)
my_pwsid='01-11-010'
ex9=lapply(X=the_years,
           FUN=scrap.it,
           the_pwsid=my_pwsid) %>% bind_rows()
ex9.plot=ggplot(ex9,aes(x=Date,
                    y=Max.Day.Use))+
  geom_line()+
  geom_smooth(method="loess",se=FALSE) +
  ylab("Max Daily Withdrawals")
print(ex9.plot)
```

>Question: Just by looking at the plot (i.e. not running statistics), does Asheville have a trend in water usage over time?
>Yes, the water usage in Asheville has a rising trend.